Supplementary material

### 2018-02-04 Manuscript/ms.tex
Initial scaffolding of discussion edits

@@

@@

As implemented, mosquito mortality is the only source of uncertainty/stochasticity in the transmission process.
As a result, the estimated mosquito mortality could be soaking up other sources of stochasticity or model misspecification.
Our implementation of the mortality process is conceptually similar to the empirical forcing functions developed by Hooker and Ellner \cite{Hooker2015}.
In that framework, Hooker and Ellner \cite{Hooker2015} estimated nonparametric functions in time that modify either the differential equations or specific model parameters to provide a good fit to the data.
These forcing functions serve as residuals on the time derivatives, and can be more readily interpreted as indicators of lack-of-fit than residuals on the state variables.
We do not employ the same explicit testing framework as \cite{Hooker2015}, but we treat the estimated mosquito mortality time series in the same spirit.
In particular, the weak periodic structure in the estimated $\epsilon_{dt}$ suggests that there may be some unmodeled, higher frequency process influencing transmission, either through the death rate itself or through another process.
Despite this, our estimated mosquito mortalities generally fell within the reasonable range from the literature \cite{Maciel-de-Freitas2008, Brady2013}.
The fact that our estimated mortality rate varied seasonally and increases with temperature above 20 degrees also agrees with the empirical literature on mosquito survival \cite{Yang2009}, though our estimated relationship is perhaps more extreme than estimated in most laboratory studies.
These observations suggest that our estimated mosquito mortality rates may actually be capturing the patterns of mosquito mortality in this system and are not obviously influenced by other sources of uncertainty or misspecification.

@@

This suggests that our model may overestimate the efficacy of mosquito control, but it also points a way toward more efficient and effective control methods.
Our results suggest that applying control early in the year is critical to disrupting city-wide disease spread, but targeting the city's entire mosquito population is likely infeasible and spatial prioritization of control is also necessary \cite{Pepin2015}.
If disease risk is driven by high levels of contact with a relatively small proportion of the mosquito population \cite{Yoon2012}, control policies that preferentially target these subsets of the mosquito population are likely to be very effective.
However, these subsets can be difficult to identify in practice.
One possible strategy is to employ contact-tracing, which can identify and target specific premises that have been directly implicated in transmission \cite{Vazquez-Prokopec2017}.
Though effective, this strategy is extremely labor-intensive and would be difficult to implement in endemic areas.

Due to these difficulties, new methods are needed that can make use of currently available surveillance data to more effectively guide spatial prioritization of control \cite{Pepin2015}.
MosquiTRAP surveillance data are available at the level of the individual trap, and case reports for the city of Vitoria are available for each of the city's 75 neighborhoods.
To interface with these data, spatial structure and human movement could be incorporated into models of dengue transmission  (e.g., through the methods outlined in \cite{Cosner2015}) but at the cost of substantial statistical and computational burden.
For instance, embedding the model presented here into a metapopulation framework based on the city's neighborhoods leads to a very large, unwieldy mechanistic model.
Moreover, without external data on human movement (possible sources of which are discussed in \cite{Stoddard2009}), it would be very difficult to identify its role in disease dynamics.
These dimensionality and identifiability issues would likely exacerbate existing difficulties inherent in fitting ODE models to data \cite{Girolami2008, Calderhead2011}.
Developing such spatial models and the methods necessary to deal with them is a promising area of future research.
In the absence of such methods, spatial targeting of control based on more traditional methods (e.g., local entomological thresholds) could be combined with temporal targeting based on city-scale mechanistic modeling.
This would allow mosquito control efforts to exploit the critical timing identified by the mechanistic modeling, while also taking advantage of the spatial richness of the surveillance data to allocate limited resources within that window.

>>>>

\subsection*{Summary}

@@

\subsection*{Model validation and interpretation of $d_v$}

The dynamics of infectious diseases, and ecological systems more broadly, are often driven by the combined efforts of intrinsic non-linearities, seasonality, and stochasticity.
We capture the interplay of these factors primarily through our specification of the mosquito mortality process.
As implemented, mosquito mortality is the only source of uncertainty/stochasticity in the transmission process.
As a result, the estimated mosquito mortality could be soaking up other sources of stochasticity or model misspecification.
Our implementation of the mortality process is conceptually similar to the empirical forcing functions developed by Hooker and Ellner \cite{Hooker2015}.
In that framework, Hooker and Ellner \cite{Hooker2015} estimated nonparametric functions in time that modify either the differential equations or specific model parameters to provide a good fit to the data.
These forcing functions serve as residuals on the time derivatives, and can be more readily interpreted as indicators of lack-of-fit than residuals on the state variables.
We do not employ the same explicit testing framework as \cite{Hooker2015}, but we treat the estimated mosquito mortality time series in the same spirit.
In particular, the weak periodic structure in the estimated $\epsilon_{dt}$ suggests that there may be some unmodeled, higher frequency process influencing transmission, either through the death rate itself or through another process.
Despite this, our estimated mosquito mortalities generally fell within the reasonable range from the literature \cite{Maciel-de-Freitas2008, Brady2013}.
The fact that our estimated mortality rate varied seasonally and increases with temperature above 20 degrees also agrees with the empirical literature on mosquito survival \cite{Yang2009}, though our estimated relationship is perhaps more extreme than estimated in most laboratory studies.
These observations suggest that our estimated mosquito mortality rates may actually be capturing the patterns of mosquito mortality in this system and are not obviously influenced by other sources of uncertainty or misspecification.

\subsection*{Implications for control policy}

@@

\subsection*{Caveats and extensions}

@@
